---
title: "Getting interest rates from State Banks"
author: "Randall Romero-Aguilar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette shows how to download time series on lending interest rates. In the first example, the user knows the BCCR _cuadro_ numbers. In the second, he uses the provided `bancarias` data frame to get those numbers.


```{r}
library(bccr)
library(ggplot2)   # to make plots
library(magrittr)  # for pipe operators
library(tidyr)     # to manipulate the data frame
library(dplyr)
library(data.table)
```

## User knows table numbers

Here, We are going to compare the local-currency interest rates on housing morgages from public (table 495) and private (table 496) banks. Those two tables contain daily data.
```{r}
vars <- list(public=495, private=496)
vivienda <- read_daily_series(vars)
head(vivienda)
tail(vivienda)
```

Now lets make a plot of monthly data from 2000 to 2015, by taking the mean of the daily observations.
```{r, cache=TRUE,fig.width=8}
vivienda <- read_daily_series(vars, 2000, 2015, 'm', mean)
head(vivienda)
tail(vivienda)


vivienda %>% 
  gather('owner','rate',2:3) %>%
  ggplot(aes(x=fecha, y=rate, color=owner)) + 
  geom_line(size=2.0) + 
  labs(title='Lending interest rate for housing, by type of bank', x='',y='% rate ')
```


## Using `bancarias` dataframe

The `bccr` package includes a `bancarias` `data.table`, which summarizes the available banking rates. In this example, we compare the lending rates of state banks by currency and activity.

First, we take the `bancarias` `data.table` to filter the desired rates and make a name ('rate') for the individual time series.
```{r}
setkey(bancarias, tipo, entidad)
myvars <- bancarias[.('Activas','Bancos Estatales'), .(rate=paste(actividad, moneda,sep = "_"), cuadro)]
```

We use the `read_daily_series` to download the data
```{r, cache=TRUE}
estatal <- read_daily_series(myvars)
```

Now let's compare the rates my loan currency, for each activity
```{r, fig.show='hold', fig.width=8, fig.height=8}
estatal %>% 
  gather("activ_moneda", "rate", 2:ncol(estatal)) %>% 
  separate(activ_moneda,c("actividad", "moneda"),"_") %>% 
  ggplot(aes(fecha, rate, color=moneda)) + 
  geom_line() + 
  facet_wrap(~actividad)
```


