---
title: "Getting interest rates from State Banks"
author: "Randall Romero-Aguilar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading interest rates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette shows how to download time series on lending interest rates. In the first example, the user knows the BCCR _cuadro_ numbers. In the second, he uses the provided `bancarias` data frame to get those numbers.


```{r, results='hide'}
library(bccr)
library(ggplot2)   # to make plots
library(magrittr)  # for pipe operators
library(tidyr)     # to manipulate the data frame
library(dplyr)
library(data.table)
```

## User knows table numbers

Here, We are going to compare the local-currency interest rates on housing morgages from public (table 495) and private (table 496) banks. Those two tables contain daily data.
```{r}
vars <- list(public=495, private=496)
vivienda <- read_daily_series(vars)
print(vivienda)
```

By default, *read_daily_series* will read all available data. To limit the sample, you can specify the first and last years to read. Moreover, we can request to change the frequency from daily to monthly by setting the option *freq='m'*.


For example, let's get monthly data from 2000 to 2015, by taking the mean of the daily observations.
```{r, cache=TRUE}
vivienda <- read_daily_series(vars, 2000, 2015, 'm', mean)
print(vivienda)
```


```{r, fig.width=8}
vivienda %>% 
  gather('owner','rate',2:3) %>%
  ggplot(aes(x=fecha, y=rate, color=owner)) + 
  geom_line(size=2.0) + 
  labs(title='Lending interest rate for housing, by type of bank', x='',y='% rate ')
```



## Using `bancarias` dataframe

The `bccr` package includes a `bancarias` `data.table`, which summarizes the available banking rates. In this example, we compare the lending rates of state banks by currency and activity.

First, we take the `bancarias` `data.table` to filter the desired rates and make a name ('rate') for the individual time series.
```{r}
setkey(bancarias, tipo, entidad)
myvars <- bancarias[.('Activas','Bancos Estatales'), .(rate=paste(actividad, moneda,sep = "_"), cuadro)]
```

We use the `read_daily_series` to download the data
```{r, cache=TRUE}
estatal <- read_daily_series(myvars)
```

Now let's compare the rates my loan currency, for each activity
```{r, fig.show='hold', fig.width=8, fig.height=8}
estatal %>% 
  gather("activ_moneda", "rate", 2:ncol(estatal)) %>% 
  separate(activ_moneda,c("actividad", "moneda"),"_") %>% 
  ggplot(aes(fecha, rate, color=moneda)) + 
  geom_line() + 
  facet_wrap(~actividad)
```


